\documentclass[a5paper]{article}

\input{header.tex}
\input{./tex-preamble/logic.tex}
\input{./tex-preamble/crypto.tex}

\begin{document}

\title{Cryptography}
\author{Langston Barrett}
\date{Fall 2017}
\maketitle
\tableofcontents
\vspace{1em}
\hrule
\vspace{1em}

\begin{itemize}
  \item Instructor: Adam Groce
  \item Textbook:
    \begin{itemize}
      \item Title: Introduction to Modern Cryptography, 2\textsuperscript{nd} Edition
      \item Author: Jonathan Katz and Yehuda Lindell
      \item ISBN: 978-1-4665-7026-9
    \end{itemize}
\end{itemize}

\section{Probability}
\label{sec:probability}

\begin{lemma}
  If
  \begin{premises}
    \item $A$ and $B$ are random variables,
    \item $B$ is sampled from some finite set of outcomes $\caB$,
  \end{premises}
  then
  \begin{conclusion}
    \begin{equation*}
      \Pr\paren{A}=\sum_{b\in \caB}\Pr\paren{A|B}\cdot\Pr\paren{B=b}
    \end{equation*}
  \end{conclusion}
\end{lemma}

\begin{lemma}[Union Bound]
  For events $E_0,\ldots,E_n$,
  \begin{equation*}
    \Pr\paren{\bigvee_{i=0}^n E_i} \leq \sum_{i=0}^n\Pr E_i
  \end{equation*}
\end{lemma}

\begin{theorem}[Bayes' Theorem]
  \begin{equation*}
    \Pr(A|B) = \frac{\Pr(B|A)\cdot \Pr A}{\Pr B}
  \end{equation*}
\end{theorem}

\section{Concepts}
\label{sec:concepts}

\begin{note}
  \begin{field}
    What's wrong with having a small key space $\caK$?
  \end{field}

  \begin{field}
    It makes your scheme vulnerable to brute-force attacks, especially when the
    distribution on the message space $\caM$ is well-understood (as in all
    adversarial experiments).
  \end{field}
\end{note}

\begin{note}
  \begin{field}
    What are the four kinds of security experiments?
  \end{field}

  \begin{field}
    \begin{enumerate}%[label=\Alph*.]
      \itemsep0em
      \item Ciphertext-only
      \item Known-plaintext
      \item Chosen-plaintext
      \item Chosen-ciphertext
    \end{enumerate}
  \end{field}
\end{note}

\begin{note}
  \begin{field}
    How does a reduction work?
  \end{field}

  \begin{field}
    In it's most general form, reduction is a tool used to show that
    problem/language $A$ is just as ``hard'' as problem/language $B$.
    \begin{enumerate}%[label=\Alph*.]
      \itemsep0em
      \item Assume that problem $B$ ``hard''.
      \item Assume $\caA$ is an algorithm that solves $A$.
      \item Using $\caA$ as a subroutine, construct a solution $\caB$ for $B$.
      \item This contradicts the assumption that $B$ couldn't be solved,
        conclude by contradiction that no such $\caA$ exists.
    \end{enumerate}
  \end{field}
\end{note}

\begin{note}
  \begin{field}
    What is one piece of information that almost every encryption scheme leaks?
    Why might it be a problem? When can and when can't it be solved?
  \end{field}

  \begin{field}
    Plaintext length. It might be a problem if
    $\caM=\braces{\text{``yes''},\text{``no''}}$. It can be solved when the
    maximum length of the encrypted messages is known in advance.
  \end{field}
\end{note}

\begin{note}
  \begin{field}
    Why is it necessary to use randomness in encryption?
  \end{field}

  \begin{field}
    No non-random scheme has indistinguishability for multiple encryptions.
  \end{field}
\end{note}

\section{Symmetric-key cryptography}
\label{sec:symmetric}

\begin{definition*}[Private-key (symmetric) encryption scheme]
  A \define{private-key encryption scheme} consists of:
  \begin{itemize}
    \itemsep0em
    \item a \define{message space} $\caM$,
    \item a \define{key space} $\caK$, and
    \item a trio of algorithms $(\Gen,\Enc,\Dec)$.
  \end{itemize}
  A scheme is \define{correct} if
  \begin{equation*}
    \apply{\Dec_k}{(\apply{\Enc_k}{m})}=m
  \end{equation*}
  for all $m\in\caM$ and $k\in\caK$.
\end{definition*}

\begin{definition*}[Perfect secrecy]
  A private-key encryption scheme $\Pi=(\Gen,\Enc,\Dec)$ is \define{perfectly
  secret} if for all distributions on $\caM$, $m\in\caM$, and $c\in\caC$
  (with $\Pr(C=c)>0$),
  \begin{equation*}
    \Pr(M=m|C=c)=\Pr(M=m)
  \end{equation*}
  Equivalently, for all $m,m'\in\caM$, and $c\in\caC$,
a
  \begin{equation*}
    \Pr(\Enc_K(m)=c)=\Pr(\Enc_K(m')=c)
  \end{equation*}
\end{definition*}

\begin{definition*}[$\mathsf{PrivK}^{\mathsf{eav}}_{\caA,\Pi}$]
  The \define{perfect adversarial indistinguishability experiment
  $\mathsf{PrivK}^{\mathsf{eav}}_{\caA,\Pi}$} is:
  \begin{enumerate}%[label=\Alph*.]
    \itemsep0em
    \item The adversary $\caA$ outputs $m_0,m_1\in\caM$.
    \item
      \begin{enumerate}%[label=\Alph*.]
        \itemsep0em
        \item A key $k\leftarrow\Gen(1^n)$ is generated.
        \item A bit $b\leftarrow \braces{0,1}$ is chosen.
        \item A ciphertext $c\leftarrow\apply{\Enc_k}{m_b}$ is fed to $\caA$.
      \end{enumerate}
    \item $\caA$ outputs $b'\in\braces{0,1}$.
  \end{enumerate}
  The experiment outputs $1$ when $b=b'$.
\end{definition*}

\begin{definition*}[Perfect indistinguishability in the presence of an eavesdropper]
	A scheme $\Pi$ has \define{perfectly indistinguishable encryptions in the
  presence of an eavesdropper} if
  \begin{equation*}
    \Pr\paren{\mathsf{PrivK}^{\mathsf{eav}}_{\caA,\Pi}=1}=\frac 12
  \end{equation*}
  for all $\caA\in\mathsf{BPT}$.
\end{definition*}

\begin{definition*}[$\mathsf{PrivK}^{\mathsf{eav}}_{\caA,\Pi}(n)$]
  The \define{adversarial indistinguishability experiment
  $\mathsf{PrivK}^{\mathsf{eav}}_{\caA,\Pi}(n)$} is:
  \begin{enumerate}%[label=\Alph*.]
    \itemsep0em
    \item The adversary $\caA$ is given $1^n$ and outputs $m_0,m_1\in\caM$ with
      $|m_0|=|m_1|$.
    \item
      \begin{enumerate}%[label=\Alph*.]
        \itemsep0em
        \item A key $k\leftarrow\Gen(1^n)$ is generated.
        \item A bit $b\leftarrow \braces{0,1}$ is chosen.
        \item A ciphertext $c\leftarrow\apply{\Enc_k}{m_b}$ is fed to $\caA$.
      \end{enumerate}
    \item $\caA$ outputs $b'\in\braces{0,1}$.
  \end{enumerate}
  The experiment outputs $1$ when $b=b'$.
\end{definition*}

\begin{definition*}[Indistinguishability in the presence of an eavesdropper]
	A scheme $\Pi$ has \define{indistinguishable encryptions in the presence of an
  eavesdropper} if
  \begin{equation*}
    \Pr\paren{\mathsf{PrivK}^{\mathsf{eav}}_{\caA,\Pi}=1}\leq \frac 12 + \apply{\negl}{n}
  \end{equation*}
  for all $\caA\in\mathsf{BPT}$ and $n\in\N$, where the probability is taken
  over randomness of $\caA$ and that of the experiment.
\end{definition*}

\begin{definition*}[$\mathsf{PrivK}^{\mathsf{mult}}_{\caA,\Pi}(n)$]
  The \define{multiple-message indistinguishability experiment
  $\mathsf{PrivK}^{\mathsf{mult}}_{\caA,\Pi}(n)$} is:
  \begin{enumerate}%[label=\Alph*.]
    \itemsep0em
    \item The adversary $\caA$ is given $1^n$ and outputs lists
      $(m_{0,0},m_{1,0},\ldots,m_{t,0})$ and $(m_{0,1},m_{1,1},\ldots,m_{t,1})$
      such that $|m_{i,0}|=|m_{i,1}|$ for all $i\in\braces{1,\ldots,t}$.
    \item
      \begin{enumerate}%[label=\Alph*.]
        \itemsep0em
        \item A key $k\leftarrow\Gen(1^n)$ is generated.
        \item A bit $b\leftarrow \braces{0,1}$ is chosen.
        \item The ciphertexts $(\apply{\Enc_k}{m_{0,b}},\ldots,\apply{\Enc_k}{m_{t,b}})$
          are given to $\caA$.
      \end{enumerate}
    \item $\caA$ outputs $b'\in\braces{0,1}$.
  \end{enumerate}
  The experiment outputs $1$ when $b=b'$.
\end{definition*}

\subsection{Peseudorandomness}
\label{subsec:pseudorandomness}

\begin{definition}
	A \textit{determinisitic} algorithm $G\in\mathsf{P}$ is a \define{pseudorandom
  generator} if there exists some real polynomial $l$ such that
  $D:\braces{0,1}^n\to\braces{0,1}^{\apply{l}{n}}$ and the following conditions hold:
  \begin{enumerate}%[label=\Alph*.]
    \itemsep0em
    \item \define{Expansion:} For all $n\in\N$, $\apply{l}{n}>n$.
    \item \define{Pseudorandomness:} For all distinguishers $D\in\mathsf{BPT}$,
      there exists a negligible function $\negl$ such that
      \begin{equation*}
        \abs{\Pr(D(r)=1)-\Pr(D(G(s))=1)}\leq \apply{\negl}{n}
      \end{equation*}
      where the first probability is taken over the choice of a uniformly random
      string $r\leftarrow\braces{0,1}^{l(n)}$ and the second over a choice of a
      uniformly random $s\leftarrow\braces{0,1}^n$, and both over randomness of
      $D$. 
  \end{enumerate}
\end{definition}

\begin{note}
  \begin{field}
    What is meant by the phrase ``let $s$ be a random string''?
  \end{field}

  \begin{field}
    Strictly speaking, this phrase doesn't make sense. A given string (or
    function) can't be \textit{random}. What it means is ``let $s$ be a string
    drawn uniformly at random from the set of all strings''.
  \end{field}
\end{note}

\begin{note}
  \begin{field}
    Does the seed of a pseudorandom generator need to be kept secret? Why?
  \end{field}

  \begin{field}
    Yes. Consider the modified one-time pad scheme where a PRG is used to expand
    the key length. If the adversary knows the seed, they know the key.
  \end{field}
\end{note}

\begin{definition}
	A \define{stream cipher} is a pair of deterministic algorithms $(\Init,\GetBits)$
  where 
  \begin{itemize}
    \itemsep0em
    \item $\Init$ takes as input a seed $s$ and an optional initialization
      vector IV, and outputs an initial state $s_0$.
    \item $\GetBits$ takes a state $s_i$ and outputs a bit $b$ and an updated
      state $s_{i+1}$.
  \end{itemize}
\end{definition}

\subsection{CPA-security}
\label{subsec:cpa-security}

\begin{definition*}[$\mathsf{PrivK}^{\mathsf{cpa}}_{\caA,\Pi}(n)$]
  The \define{CPA indistinguishability experiment
  $\mathsf{PrivK}^{\mathsf{cpa}}_{\caA,\Pi}(n)$} is:
  \begin{enumerate}%[label=\Alph*.]
    \itemsep0em
    \item A key $k\leftarrow\Gen(1^n)$ is generated.
    \item The adversary $\caA$ is given $1^n$ and access to the oracle
      $\Enc_k(-)$. The adversary outputs $m_0,m_1\in\caM$ with $|m_0|=|m_1|$.
    \item
      \begin{enumerate}%[label=\Alph*.]
        \itemsep0em
        \item A bit $b\leftarrow \braces{0,1}$ is chosen.
        \item The ciphertext $c\leftarrow\apply{\Enc_k}{m_b}$ is fed to $\caA$.
      \end{enumerate}
    \item $\caA$ continues to have access to $\Enc_k(-)$ and outputs
      $b'\in\braces{0,1}$.
  \end{enumerate}
  The experiment outputs $1$ when $b=b'$.
\end{definition*}

\end{document}
